---
title: "Discrimination"
author: "David Kane"
format: html
execute: 
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
```

## From Monday's Class

### Background Information

Bertrand, Marianne and Sendhil Mullainathan. 2004. "Are Emily and Greg More Employable Than Lakisha and Jamal? A Field Experiment on Labor Market Discrimination." American Economic Review, 94 (4): 991–1013. Data cleaned and discussed in "Quantitative Social Science: An Introduction" by Kosuke Imai.

"We study race in the labor market by sending fictitious resumes to help-wanted ads in Boston and Chicago newspapers. To manipulate perceived race, resumes are randomly assigned African-American- or White-sounding names. White names receive 50 percent more callbacks for interviews. Callbacks are also more respon- sive to resume quality for White names than for African-American ones. The racial gap is uniform across occupation, industry, and employer size. We also find little evidence that employers are inferring social class from the names. Differential treatment by race still appears to still be prominent in the U.S. labor market."

### Discussion

Subject: Discrimination

Broad question: Race and employment in the United States

Specific question: Effect of racially-coded names on call-backs

Preceptor Table: 
  
  * Units: individuals (or resumes ?) 
  
  * (Potential) Outcomes: call-back if black name, call-back if white name   
  
  * Causal/predictive: a causal model  
  
  * Covariates: sex, zip code (not shown in table)  
  
  * Treatment: black name on resume versus white name on resume 

## Example Preceptor Table
  
```{r}
tibble(ID = c("1", "2", "...", "10", "11", "...", "N"),
       attitude_after_control = c("Yes*", "No", "...", "No*", "Yes*", "...", "Yes"),
       attitude_after_treated = c("Yes", "Yes*", "...", "No", "Yes", "...", "No*"),
       treatment = c("Black", "White", "...", "Black", "Black", "...", "White")) |>
  
  gt() |>
  tab_header(title = "Preceptor Table") |> 
  cols_label(ID = md("ID"),
             attitude_after_control = md("Callback if White Name"),
             attitude_after_treated = md("Callback if Black Name"),
             treatment = md("Treatment")) |>
  tab_style(cell_borders(sides = "right"),
            location = cells_body(columns = c(ID))) |>
  tab_style(style = cell_text(align = "left", v_align = "middle", size = "large"), 
            locations = cells_column_labels(columns = c(ID))) |>
  cols_align(align = "center", columns = everything()) |>
  cols_align(align = "left", columns = c(ID)) |>
  fmt_markdown(columns = everything()) |>
  tab_spanner(label = "Potential Outcomes", columns = c(attitude_after_control, attitude_after_treated)) |>
  tab_spanner(label = "Covariate", columns = c(treatment))
```

## Questions for Second Class

<!-- There are no "right" answers. Each person/group might have a different goal . . . -->

> What are the units, precisely?

The resumes tied to the individuals.

> What is the moment in time, precisely?

The moment the resumes are reviewed, or when the study was taken; 2004. The most relevant time period would be the current moment - today.

> What is one reason why validity might not hold?

Racial profiling and bias in the one reviewing the resumes.

<!-- Assume validity holds, and we can create the Population Table -->

> Describe the Population Table in words

We may also add the Name's of the units, which would complement the ID column. Then we could also add the "sex" column, which would be, presumably, a covariate. 

> What is one reason why stability might not hold?

Major world events could have taken place between the date of the study to current date which would affect the correspondence between the data.  

> What is one reason why representativeness might not hold?

Lack of diversity within fields of companies and companies themselves. Does the data represent the rest of the country or only a specific location? -> In our case, it only affects Boston, Chicago, so the representativeness would not hold in this scenario.

> What is one reason why unconfoundedness might not hold?

The entropical sense and the value of the ambiguity and randomness of the treatment values associated with the data.

> Describe in words what the Preceptor Table would look like if we were trying to, instead, predict whether or not someone would get a callback?

We would not have the treatment column, and the data would we a lot more oriented towards being predictive towards the Nth ID.

<!-- If time, load up and explore the data.  -->
  